---
title: "DATASET_REPORT"
author: "Ingenieros del Futuro"
date: "20 de octubre de 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries
```{r libraries, include=TRUE}
library(tidyverse)
library(readxl)
library(lubridate)
library(caret)
library(stats)
```
# Dataset location
```{r datalocation, include=TRUE}
dataset_raw = read.csv("ldex_20161118/data_calibrated/mass/DATASET.tab",
                       sep = " ")
```

# Tidy data
The next step was to tidy the dataset.
For the extension of code, this would not be included but is located in the RScript_2.R
```{r tidy, include=FALSE}
dataset_raw = dataset_raw[,c(1,2,9)]
dataset_raw = set_names(dataset_raw, nm = c('time-stamp', 'value', 'saturated'))

dataset_raw = dataset_raw %>% 
  separate(`time-stamp`, into = c("year", "hour"), 
           sep = "-")
dataset_raw = dataset_raw %>% 
  separate(hour, into = c("day", "time"),
           sep = "T")

dataset_raw
dataset_raw = dataset_raw %>% 
  filter(saturated == 0)

dataset13 = dataset_raw %>% 
  filter(year == "2013") %>% 
  mutate(day = as.integer(day)) %>% 
  mutate(day = as.Date(day, "2012-12-31"))

dataset14 = dataset_raw %>% 
  filter(year == "2014") %>% 
  mutate(day = as.integer(day)) %>% 
  mutate(day = as.Date(day, "2013-12-31"))

dataset_tidy = bind_rows(dataset13, dataset14)

sum_dat = dataset_tidy %>% 
  group_by(day) %>% 
  summarize(mean(value))
```
# Data Plot
For data analysis we would make a plot. In this plot we observe a cyclical behavior of the data
```{r plot}
sum_dat %>% 
  ggplot(aes(day, `mean(value)`)) +
  geom_line()
```

# Fit a model prediction
Next step was to fit a prediction model for the dataset in order to be able to predict future or not given data points 
```{r fit}
p <- 0.8
set.seed(1)

test_index = sample.int(n = nrow(sum_dat),
                        size = floor(p*nrow(sum_dat)),
                        replace = FALSE)
train = sum_dat[test_index,]
test = sum_dat[-test_index,]

model = lm(`mean(value)` ~ day, data = train)
valuepred = predict(model, test)

summary(model)

actual_preds <- data.frame(cbind(actuals = test$`mean(value)`,
                                 predicted = valuepred))
actual_preds['error'] = actual_preds$actuals - actual_preds$predicted
actual_preds %>% 
  summarize(RMSE = (sum(error)^2)/n())
```
We conclude that datapoints are insufficient to fit a prediction model to the dataset. Nevertheless due to the cyclical behaviour we suggest a Exponential Smoothing algorithm with 3 year data collection
